---
title: Trade Analysis with R
author: Admin
date: '2019-08-05'
slug: trade-analysis-with-r
categories: []
tags: 
  - R
  - Trade Analysis
  - Economics
summary: 'A Quick Intro into the comtradr Package'
lastmod: '2019-08-05T15:24:38+02:00'
featured: no
projects: []
---
In order to access trade information, we'll use the [comtradr](https://cran.r-project.org/web/packages/comtradr/comtradr.pdf) package which provides an API access to the [UN Comtrade Database](https://comtrade.un.org/).

```{r eval= TRUE, message=FALSE, warning=FALSE, results='hide'}
library(comtradr)
```

Suppose we want to analyze the Algerian exports' evolution (Montary values of goods and services) in relation to the Ouest-European market which represents an important one. 
To that aim, we'll use the **ct_search** function. 

```{r eval = TRUE}

export_DZ <- ct_search(reporters = "Algeria", 
  partners = c("France", "Spain", "Italy", "Germany"), 
  trade_direction = "exports")

#Note that the API limitS our request to 5 partner. For more information on API limits, refer to the the comtradr documentation.  


```

Let us embellish our database and select the relevant information

```{r message=FALSE, warning=FALSE, echo=TRUE, results='markup'}

library(tidyverse)

export_DZ <- as_tibble(export_DZ)

export_DZ <- export_DZ %>% select(year, partner, trade_value_usd)

head(export_DZ)


```

Finally, let's plot our data to have a broader overview

```{r message=FALSE, warning=FALSE, fig.show='asis', fig.width = 10}

options(scipen = 999) # avoiding scientific values (e.g. 10e9)

ggplot(export_DZ, aes(x = year, y= ((trade_value_usd)/1000))) +  
  facet_grid( ~ as.factor(partner))+                                     
  geom_line(size = 2, color = "lightgreen") + 
  ylab("trade value in 1000$")+
  xlab("years")+
  scale_x_continuous(breaks = c(seq(1992, 2015, by = 2), 2017))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

Interesting to see that the export value towards Germany is relatively low. Let us compare the exports to the imports. 


```{r eval = TRUE}

import_DZ <- ct_search(reporters = "Algeria", 
  partners = c("France", "Spain", "Italy", "Germany"), 
  trade_direction = "imports")

```

We'll use the [patchwork](https://github.com/thomasp85/patchwork) package to combine the import and the export plots. It's as easy as drinking from a bottle of water. 


```{r message=FALSE, warning=FALSE}

library(patchwork)

max <- max(import_DZ$trade_value_usd/1000, export_DZ$trade_value_usd/1000)

min <- min(import_DZ$trade_value_usd/1000, export_DZ$trade_value_usd/1000)


p1 <- ggplot(export_DZ, aes(x = year, y= ((trade_value_usd)/1000))) +  
  facet_grid( ~ as.factor(partner))+                                     
  geom_line(color = "#FF5E00") + 
  ylab("exports value in 1000$")+
  xlab("years")+
  scale_x_continuous(breaks = c(seq(1992, 2015, by = 4), 2017))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))


p2 <- ggplot(import_DZ, aes(x = year, y= ((trade_value_usd)/1000))) +  
  facet_grid( ~ as.factor(partner))+                                     
  geom_line(color = "#FF00A2") + 
  ylab("imports value in 1000$")+
  xlab("years")+
  scale_x_continuous(breaks = c(seq(1992, 2015, by = 4), 2017))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))



p1 + p2 + patchwork::plot_layout(ncol = 1) + ylim(min, max)

# ylim(min, max) allows us to have the same dimensiality in our two plots.  


```
















