---
title: 'Introduction to Deep Learning, What is a Tensor ? '
author: admin
date: '2019-08-09'
tags:
  - Deep Learning
  - R
  - Tensorflow
  - Tensors
slug: introduction-to-deep-learning-what-is-a-tensor
lastmod: '2019-08-09T22:07:19+02:00'
featured: no
---

# Tensors Definition

Ever wandered why one of the most used Deep Learning library, **Tenserflow** is called that way ? the name came from the fundamentals **Tensors**. 

According to Chollet F and Allaire JJ (Deep Learning with R und Keras, 2018), Tensors are a generalization of vectors and matrices to an arbitrary number of dimensions. 

## 0-Dimension Tensors

Scalars are considered to be the simplest form of Tensors. 


```{r}
zero_Dim_Tens <- 6
zero_Dim_Tens
```


## 1-Dimension Tensors

Vectors are 1-Dimension Tensors. 

```{r}

one_Dim_Tens <- c(12, 3, 6, 14, 10)
one_Dim_Tens

str(one_Dim_Tens)
```


## 2-Dimensions Tensors

Matrices are 2-Dimensions Tensors. 

```{r}

two_Dim_Tens <- matrix(rep(1, 3*3), nrow = 3, ncol = 3)
two_Dim_Tens

str(two_Dim_Tens)


```


## 3 and higher-dimensional Tensors

*array* objects are Tensors which can have more than 2 dimensions. For example, if we include the above matrice into a new array: 


```{r}

two_Dim_Tens <- array(rep(1, 3*3), dim = c(2, 3, 2))
two_Dim_Tens

str(two_Dim_Tens)
```

A Tensor whith the propoerty *num [1:2, 1:3, 1:2]* can be read as an Integer Tensor with three *axes* (Dimensions) and a shape of *2 3 2*.




