---
title: 'An Introduction to the ggplot2 package. '
author: admin
date: '2019-08-08'
slug: an-introduction-to-the-ggplot2-package
categories: []
tags: []
subtitle: ''
summary: ''
authors: []
lastmod: '2019-08-08T01:12:03+02:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---


The [ggplot2](https://ggplot2.tidyverse.org/) package integrates the **tidy** philosophy of programming introduced mainly by [Hadley Wickham](https://twitter.com/hadleywickham). It is part of the **tidyverse** package which includes several other packages related to the tidy world. Through this is tutorial, we'll present some basic ggplot functionalities using the **mtcars** data base. 

First of of all, you need to install the ggplot2 package which can be done either by 

```{r eval = FALSE}
install.packages("tidyverse")
```

or 

```{r eval = FALSE}
install.packages("ggplot2")
```

# Plotting with ggplot

The mtcars object is a dataframe included in R that displays 10 aspects of automobile design and performance for 32 automobiles (1973–74 models). Let us have a quick overview of the mtcars dataset. 


```{r}

head(mtcars)

```


We want to plot the relation between the miles per galon *(mpg)* and the weights *(wt)* of our cars' models. To do that in ggplot2 one need to identify three principal elements (or two depending on your desired graph): The *database* from which the information will be extracted, the *x-axis variable* and the *y-axis variable*. 

```{r}
library(ggplot2) # or library(tidyverse)

ggplot(data = mtcars, mapping = aes(x = mpg, y = wt))

```

In the exemple above, we used the function **ggplot** and two parameters *data * and *mapping* to structure our graph. The *mapping* parameters determine which variable should be considered in which axis. To plot our volue we need to add another argument to our code. Suppose we want to plot a scatterplot : 

```{r}

ggplot(data = mtcars, mapping = aes(x = mpg, y = wt)) +
  geom_point()   # geom_point() refers to scatterplot

```

To modify the size of the color of the displayed points, we just introduce the following parameters in the geom_point() function :  


```{r}

ggplot(data = mtcars, mapping = aes(x = mpg, y = wt)) +
  geom_point(size = 3, color = "red")   # geom_point() refers to scatterplot

```


For an exhaustive list of R colours, just check this [website](http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf).\n
Now, if we want to modify the theme of our plot, proceed as follow : 

```{r}

ggplot(data = mtcars, mapping = aes(x = mpg, y = wt)) +
  geom_point(size = 3, color = "red") + 
  theme_bw()  #just tape theme and wait to see several choices

```



finally, it is possible to assign a plot to a defined variable : 


```{r}

plot1 <- ggplot(data = mtcars, mapping = aes(x = mpg, y = wt)) +
  geom_point(size = 3, color = "red") + 
  theme_bw()  

```



# Plotting Interactively with Plotly


Plotly is an interesting package that allows us to create interactive web graphics from ggplot2. Just convert our above scatterplot into an interactive one, just use the function **ggplotly** : 

```{r message=FALSE}

library(plotly)  # Load the package after installing it

ggplotly(plot1)
```


Several functionalities appear in the top-right of the plot.

## Plotting Distributions with Histograms

Histogram are suitable for plotting a continous variable. If you want to make a histogram use **geom_histogram**:

```{r}

ggplot(data = mtcars, mapping = aes(x = mpg)) +
  geom_histogram()


```



We can change the size of the bins using the *binwidth* parameter : 


```{r}

ggplot(data = mtcars, mapping = aes(x = mpg)) +
  geom_histogram(binwidth = 3)


```


By default, the data is grouped into 30 bins. To can change the number of bins, we use the parameter **bins**:


```{r}

ggplot(data = mtcars, mapping = aes(x = mpg)) +
  geom_histogram(bins = 6)


```


To modify the color of the histogram, just proceed as follows: 


```{r message=FALSE, warning=FALSE}

ggplot(data = mtcars, mapping = aes(x = mpg)) +
  geom_histogram(fill = "pink", color = "yellow") # Mind the difference


```



We can also use the plotly package to plot an interactive histogram. 


```{r message=FALSE, warning=FALSE}

histogram <- ggplot(data = mtcars, mapping = aes(x = mpg)) +
  geom_histogram(fill = "pink", color = "yellow") 

ggplotly(histogram)

```

# Filling Histograms with Categories 

Suppose we are interested in ploting the distribution of the miles per gallon (mpg) variable but this time we want to differentiate between automatic and manuel cars (am : 0 = automatic, 1 = manual). First of all, we must convert the *am* column into a factor variable:


```{r message=FALSE, warning=FALSE}


mtcars$am <- factor(mtcars$am, labels = c("Automatic", "Manual")) # We provide labels 

ggplot(data = mtcars, mapping = aes(x = mpg, fill = am)) +
  geom_histogram() 

```


From the above histogram, we observe that manual vehicules have a higher *mpg* performance. 



# Splitting Plots with Facets


We show above how to plot a distribution according to a specific category. One may proceed otherwise and generate two distinc plots using **Face_grid()** : 


```{r message=FALSE, warning=FALSE}

ggplot(data = mtcars, mapping = aes(x = mpg)) +
  geom_histogram(fill = "yellow", color = "pink") +
  facet_grid(~am) +
  theme_light()


```











