---
title: Introduction to the Subset function
author: Mohamed El Fodil Ihaddaden
date: '2019-11-11'
slug: introduction-to-the-subset-function
categories: []
tags:
  - base R
  - R
subtitle: ''
summary: ''
authors: []
lastmod: '2019-11-11T14:02:12+01:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---


Although I'm a big fan of the tidyverse philosophy of handling and wrangling data, one must admit that there are quit powerful functions in base R. One of theses functions is `subset()` which returns a dataframe according to some defined subsetting properties. Let's dive into one example using the simple **mtcars** data:

```{r}

head(mtcars)  # A quick look at the mtcars data

```

Suppose, we want to extract all the vehicules that have an *mpg* greater than 20:

```{r}

subset(mtcars, mpg > 20)  # the first argument is the dataframe


```


You can observe that the *subset* function is flexible. We don't have to specify the column name with the dollar sign (*mtcars$mpg*).  

Let's take another more complex example. We will extract all vehicules that have an *mpg* superior to 30 and a *cyl* equal to 4:


```{r}

subset(mtcars, mpg > 30 & cyl == 4) # & <=> AND

```

As you can see, the *subset* function works smoothly with the R logical expressions. 

Finally, there is another important argument that we must see. Suppose, we want to extract a specified column, let's say that we're solely interested in the *wt* variable. We can extract this column depending using the `select` argument: 

```{r}


subset(mtcars, mpg > 30 & cyl == 4, select = wt)


```

In the same way, we can extract several columns : 


```{r}


subset(mtcars, mpg > 30 & cyl == 4, select = c(wt, disp, am))


```



