---
title: It's a Bird, It's a Plane ... It's a ggforce function
author: ''
date: '2020-09-15'
slug: it-s-a-bird-it-s-a-plane-it-s-a-ggforce-function
categories: []
tags:
  - ggplot2
  - tidyverse
subtitle: ''
summary: ''
authors: []
lastmod: '2020-09-15T23:53:45+02:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---


<style>

@import url('https://fonts.googleapis.com/css2?family=Noto+Sans+JP:wght@700&display=swap');


</style>





```{css, echo = FALSE}

body{

  color:#235367;
    
  font-family: 'Noto Sans JP', sans-serif;
}

img{
  border-radius: 8px;
  width=500px;
  height=500px;
}


pre{
  
  background-color:#BCEDDD;
  
  
}


```


[ggforce](https://github.com/thomasp85/ggforce) is a powerful package that extends heavily the ggplot2 API. In this blog post, I won't go through all its capabilities, I'll just provide a sample of some cool functions that can be used by intermediate ggplot2 user (like me, well I hope to) to upgrade his plotting portfolio.



<br> 
---


# theme_no_axes()

This function removes the axis text, title, ticks and grid lines for any themed plot, particularly useful in many cases:  


```{r, message=FALSE, warning=FALSE, fig.align='center'}
library(tidyverse)
library(ggforce)

ggplot(mpg, aes(hwy, cty)) +
  geom_point() +
  theme_no_axes(base.theme = theme_dark()) 

```



```{r, message=FALSE, warning=FALSE, fig.align='center'}

ggplot(mpg, aes(hwy, cty)) +
  geom_point() +
  theme_no_axes(base.theme = ggthemes::theme_economist_white()) 

```

<br>


# facet_zoom()


`facet_zoom()` zooms on a particular data category within a plot. Below we'll use the `starwars` data frame and zoom on the x-axis of the _female_ category from the _sex_ variable. 

```{r, fig.align='center'}



ggplot(starwars, aes(mass, height, col = sex)) + 
  geom_point(na.rm = TRUE) +
  facet_zoom(x = sex == "female", zoom.size = 2) +
  ggthemes::theme_hc()


```


# facet_row()

`facet_row()` is a one-dimensional version of `facet_wrap()`. It aligns all the subplot into one row:


```{r, fig.align='center'}

library(palmerpenguins)

ggplot(penguins) +
  geom_point(aes(bill_length_mm, body_mass_g), col = "white", na.rm = TRUE) +
  facet_row(~island) +
  theme_no_axes(base.theme = theme_dark()) 


```

`facet_col` does the same thing except that it arranges sub-plots within one column : 


```{r, fig.align='center'}

ggplot(penguins) +
  geom_point(aes(bill_length_mm, body_mass_g), col = "white", na.rm = TRUE) +
  facet_col(~island) +
  theme_no_axes(base.theme = theme_dark()) 


```


# facet_matrix()

Using `facet_matrix()` you can specify a matrix composed of inter-related plots. In the example below, we'll map `bill_length_mm`, `bill_depth_mm` and `flipper_length_mm` (on the x-axis) with  `body_mass_g`  (on the y-axis):

```{r, fig.align='center'}

ggplot(penguins, aes(x = .panel_x, y = .panel_y, col = sex)) +
  geom_point(na.rm = TRUE) +
  facet_matrix(rows = vars(body_mass_g), 
               cols = vars(bill_length_mm, bill_depth_mm, flipper_length_mm), 
               switch = "x")
```

You can remove the `x = .panel_x, y = .panel_y` part using the `geom_autopoint()` function: 



```{r, fig.align='center'}

ggplot(penguins, aes(col = sex)) +
  geom_autopoint(na.rm = TRUE) +
  facet_matrix(rows = vars(body_mass_g), 
               cols = vars(bill_length_mm, bill_depth_mm, flipper_length_mm), 
               switch = "x")
```

In the same way, you can get an overview of the different interactions between your numerical variables: 


```{r, fig.align='center'}

ggplot(penguins, aes(col = sex)) +
  geom_autopoint(na.rm = TRUE) +
  facet_matrix(rows = vars(bill_length_mm:body_mass_g),
               switch = "x")


```



It's also possible to include interactions between numerical and categorical variables (be sure to drop missing values beforehand)



```{r, fig.align='center', fig.height=12, fig.width=12}

ggplot(penguins %>% drop_na(), aes(col = sex)) +
  geom_autopoint() +
  facet_matrix(rows = vars(island:body_mass_g),
               switch = "x") +
  theme(axis.text.x = element_text(angle = 90))


```



Further, it's possible to plot histograms within the diagonal of the matrix using the `geom_autohistogram()` function: 

```{r, fig.align='center', fig.height=12, fig.width=12}

ggplot(penguins %>% drop_na(), aes(col = sex, fill = sex)) +
  geom_autopoint() +
  facet_matrix(rows = vars(island:body_mass_g),
               switch = "x", layer.diag = 2) +
  theme(axis.text.x = element_text(angle = 90)) +
  geom_autohistogram()

```

Here the `layer.diag` argument determines which layer will be used within the diagonal. In this case, we've specified the value 2 which corresponds to the second layer: `geom_autohistogram()`.

In the same manner, we can use `geom_autodensity()` for densities.


```{r, fig.align='center', fig.height=12, fig.width=12}

ggplot(penguins %>% drop_na(), aes(col = sex, fill = sex)) +
  geom_autopoint() +
  facet_matrix(rows = vars(island:body_mass_g),
               switch = "x", layer.diag = 2) +
  theme(axis.text.x = element_text(angle = 90)) +
  geom_autodensity()

```


# geom_mark_ellipse() 


First, we render the following plot: 

```{r, fig.align='center'}
ggplot(penguins %>% drop_na(), aes(bill_depth_mm, body_mass_g, col = species)) +
  geom_point(na.rm = TRUE)

```

Now, let's say we want to highlight a particular species (e.g. Gentoo), we'll use the `geom_mark_ellipse()` as follows: 


```{r}

ggplot(penguins %>% drop_na(), aes(bill_depth_mm, body_mass_g)) +
  geom_point() +
  geom_mark_ellipse(aes(filter = species == "Gentoo", 
                        label = "Gentoo", 
                        description = "Gentoo penguins seem quite different from the other species"),
                        label.fill = "pink",
                        label.colour = "red", 
                        con.colour = "red",
                        con.linetype = 2, 
                        expand = unit(2, "mm")) 


```



Note that there other `geom_mark_*()` functions (e.g. geom_mark_rect) but these 




```{r}
ggplot(penguins %>% drop_na(), aes(bill_depth_mm, body_mass_g)) +
  geom_point() +
  geom_mark_rect(aes(filter = species == "Gentoo", 
                        label = "Gentoo", 
                        description = "Gentoo penguins seem quite different from the other species"),
                        label.fill = "pink",
                        label.colour = "red", 
                        con.colour = "red",
                        con.linetype = 2, 
                        expand = unit(0.01, "mm"), 
                        fill = "pink") 

```
















