---
title: Introducing the flipdownr package. Create a CountDown in Rmarkdown Documents and Shiny Apps
author: ''
date: '2020-10-16'
slug: introducing-the-flipdownr-package-create-a-countdown-in-rmarkdown-documents-and-shiny-apps
categories: []
tags: []
subtitle: ''
summary: 'Introducing the flipdownr package. Create a CountDown in Rmarkdown Documents and Shiny Apps'
authors: []
lastmod: '2020-10-16T15:56:37+02:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---


```{css, echo = FALSE}

body{
  
  background-color: #BD8571;
  
}

```

# flipdownr


The goal of flipdownr is to create a Countdown object within your RMarkdown documents and Shiny application. `flipdownr` is an R wrapper of the JavaScript library [flipdown](https://pbutcher.uk/flipdown/)

## Installation

You can install the development version of flipdownr from [GITHUB](https://github.com/feddelegrand7/flipdownr) with:

```{r, eval=FALSE}

# install.packages("remotes")

remotes::install_github("feddelegrand7/flipdownr")

```

## Functions and Arguments

`flipdownr` has one function: `flipdown` which contains the following arguments: 

<hr>

+ __downto__ : the targeted date for the countdown. It must have the form of 'yyyy-mm-dd' or 'yyyy-mm-dd hh:mm:ss`. Note that you can include the time zone (example UTC): 'yyyy-mm-dd hh:mm:ss UTC';

+ __id__ : the id of the countdown. Useful when you need to include many countdown objects in one document;

+ __theme__ : The theme of the flipdown object. Two possible choices: 'dark' or 'light'. Defaults to 'dark'

+ __headings__ : A character vector pecifying the heading related to each time part. Defaults to English headings: c("Days", "Hours", "Minutes", "Seconds").


## Examples: 

This article was written the 2020-10-16 (16th of October 2020) at 16:13 (GMT+1) approximately, as such the CountDown will reflect it. 


```{r}

library(flipdownr)

flipdown(downto = "2020-11-16", id = "c1", theme = "dark")



```
<br>


You can add another level of precision by specifying the exact time: 

```{r}

flipdown(downto = "2020-11-16 22:30:00 UTC", id = "c2", theme = "dark")

```





<br>

You can also specify a `light` theme: 


```{r}

flipdown(downto = "2020-11-16", id = "clight", theme = "light")

```

<br>

You can modify the headers with the `headings` argument: 


```{r}


flipdown(downto = "2020-11-16",
         id = "cspanish", 
         theme = "light", 
         headings = c("Dias", "Horas", "Minutos", "Segundos"))


```



## Code of Conduct

Please note that the website_iha project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.

