---
title: 'fabricerin: A Tutorial'
author: Mohamed El Fodil Ihaddaden
date: '2020-08-02'
slug: fabricerin-a-tutorial
categories: []
tags:
  - fabricerin
subtitle: ''
summary: ''
authors: []
lastmod: '2020-08-02T11:15:09+02:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---




## 1. fabric_drawing()
***


> Create a canvas in which the user can draw. 



```{r}

library(fabricerin)

fabric_drawing(cid = "canvas1", 
               cwidth = 400, 
               cheight = 400, 
               cfill = "orange", 
               drawingWidth = 20, 
               gumSize = 10)

```


You can have many canvas working independently within one document. 

```{r}

fabric_drawing(cid = "canvas2", 
               cwidth = 200, 
               cheight = 200, 
               cfill = "lightblue", 
               drawingWidth = 10, 
               gumSize = 40)


```


## 2. fabric_shape()
***

> Creates a canvas in which a defined shape element is included. The shape is positioned according to two elements `left` and `top`. Both are set relatively to the canvas.

### 2.1. Rectangle 
***

```{r}


fabric_shape(cid = "canvas_Rect", 
             cwidth = 500, 
             cheight = 500, 
             cfill = "lightgreen", 
             shapeId = "rect1", 
             shape = "Rect", 
             left = 200, 
             top = 200)

```

The user can modify the size, position and rotation of the shape interactively, if you want to disable this feature set the `selectable` argument to `FALSE`: 


```{r}


fabric_shape(cid = "canvas_Rect2", 
             cwidth = 500, 
             cheight = 500, 
             cfill = "lightgreen", 
             shapeId = "rect2", 
             shape = "Rect", 
             left = 200, 
             top = 200, 
             selectable = FALSE)

```

### 2.2. Circle

```{r}


fabric_shape(cid = "canvas_circle", 
             cwidth = 500, 
             cheight = 500, 
             cfill = "red", 
             shapeId = "circ", 
             shape = "Circle",
             fill = "yellow",
             radius = 100,
             strokecolor = "green", 
             strokewidth = 2)


```


### 2.3. Triangle

```{r}


fabric_shape(cid = "canvas_Triangle", 
             cwidth = 500, 
             cheight = 500, 
             cfill = "#B4B4D8", 
             shapeId = "tri", 
             shape = "Triangle", 
             left = 150, 
             top = 150, 
             fill = "#D7BC66")

```


### 2.4. Polygon


```{r}



fabric_shape(cid = "canvas_Polygon",
             cwidth = 400, 
             cheight = 400, 
             shapeId = "poly1", 
             shape = "Polygon", 
             top = 200, 
             cfill = "#4E5C68", 
             fill = "#1BDB99",
             xPolygon = c(200, 250, 250, 150, 150), 
             yPolygon = c(10, 50, 180, 180, 50))



```




## 3. fabric_text()
***

> Creates a canvas in which a defined text is rendered. 


```{r}

fabric_text(cid = "canvas_text", 
            cwidth = 800, 
            cheight = 100, 
            cfill = "#D769AB", 
            textId = "txt1", 
            text = "You understand that you must wear a mask ?!!", 
            left = 10, 
            top = 10, 
            fill = "blue",
            strokewidth = 0)

```


## 4. fabric_image()
***

> Inserts external images within a canvas element.


```{r}

fabric_image(cid = "cimage", 
             cfill = "#000000", 
             cwidth = 500, 
             cheight = 500,
             imgId = "dplyrhex",
             imgsrc = "https://raw.githubusercontent.com/rstudio/hex-stickers/master/PNG/dplyr.png",
             imgwidth = 250,
             imgheight = 250)
                  


```


## 5. fabric_curtail()
***


> Add a background or an overlay image to a preexisting canvas. 



```{r}

fabric_shape(cid = "canvas_curtail",
              shapeId = "tri1",
              shape = "Triangle",
              fill = "lightblue")

fabric_curtail(cid = "canvas_curtail",
imgsrc = "https://upload.wikimedia.org/wikipedia/commons/thumb/1/1b/R_logo.svg/724px-R_logo.svg.png",
type = "background"

              )


```

