---
title: 'Taking notes in RMarkdown using the fabricerin package '
author: Mohamed El Fodil
date: '2020-07-26'
slug: taking-notes-in-rmarkdown-using-the-fabricerin-package
categories: []
tags: []
subtitle: ''
summary: ''
authors: []
lastmod: '2020-07-26T22:19:26+02:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---



I've recently released a new package called `fabricerin`, now you can insert easily a canvas withing RMarkdown documents and Shiny apps for drawing or notes taking. 

How to do that ? two functions: 

- `use_fabric()`: this function is mandatory and has to be set at the beginning of your document (app). It enables the javascript library [fabricjs](http://fabricjs.com/). Of course, __you'll need to run it only once.__

- `fabric_drawing()`: allows you implement a canvas in which you can draw/take notes.


## __Try it yourself, take some notes !__


```{r}

library(fabricerin)

use_fabric()

fabric_drawing(cid = "canvasId", #canvas Id 
               cwidth = 400, #canvas width
               cheight = 400, #canvas height
               cfill = "#282A36", #canvas color
               drawingWidth = 10, #width of the drawing lines
               gumSize = 30 #size of the gum
               )


```


You can change play with the arguments to shape you canvas and your drawing experience.



```{r}

fabric_drawing(cid = "canvasId2", #canvas Id 
               cwidth = 500, #canvas width
               cheight = 1000, #canvas height
               cfill = "#FFCD43", #canvas color
               drawingWidth = 5, #width of the drawing lines
               gumSize = 50 #size of the gum
               )

```






For more info, here is the [Github repo](https://github.com/feddelegrand7/fabricerin)























