---
title: R Base Gems
author: ''
date: '2020-09-10'
slug: r-base-gems
categories: []
tags:
  - base R
subtitle: ''
summary: 'Some Gems from R Base'
authors: []
lastmod: '2020-09-10T11:48:24+02:00'
featured: no
image: 
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

<style>

@import url('https://fonts.googleapis.com/css2?family=Lora&display=swap');


</style>





```{css, echo = FALSE}

body{
  background-color:#E7EBEC;
    
  font-family: 'Lora', serif;
}

img{
  border-radius: 8px;
  width=500px;
  height=500px;
}



```




# `agrep()`


<center> 

![from Unsplash by Wolfgang Hasselmann](https://images.unsplash.com/photo-1577736164649-18f4ade579d6?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=750&q=80)

</center>

The `countries` vector below lists some African countries. The last three values contain mistakes (__Algerria, Morocoo and algeri__). In real life, it's usual to work with imperfect data. The `agrep()` function allows us to deal with this specific situation by looking at __approximate__ patterns. Suppose, we want to extract the elements that contains the word __Algeria__ within the `countries` vector:

```{r}

countries <- c("Algeria", "Morocco", "Tunisia", "Mali", "Tchad", "Kenya", "Algerria", "Morocoo", "algeri")

indexes <- agrep(pattern = "Algeria", x = countries, ignore.case = TRUE)

countries[indexes]

```





# `abbreviate()`

<center> 

![from Unsplash by Kirill Pershin](https://images.unsplash.com/photo-1551944325-776c9772d382?ixlib=rb-1.2.1&auto=format&fit=crop&w=401&q=80)

</center>


The above problem can also be handled using the `abbreviate()` function: 


```{r}

# Transform the words to lower cases 

countries_lower <- tolower(countries)

abbreviate(
  
  names.arg = countries_lower, 
  
  minlength = 3, 
  
  strict = TRUE, # We permit duplications
  
  named = FALSE
)

```



# `table()` 

<center>

![from Unsplash by Marcus Spiske](https://images.unsplash.com/photo-1547497348-42767a587639?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=750&q=80)

</center>




`table()` is a famous function that displays the counts of appearance of each value within a vector. 
```{r}

countries <- c("Algeria", "Algeria", "Mali", "Kenya", "Mali", "Mali", "Senegal", "Uganda", "Senegal", "Morocco", "Senegal", "Senegal", "Senegal", NA, NA, NA, NA, NA, NA)

table(countries, useNA = "no")

```

We can change the `useNA` argument to `"always"` to get the count of NAs: 

```{r}

table(countries, useNA = "always")


```

If you want to sort by count of appearance: 


```{r}

my_tab <- table(countries, useNA = "no")

sort(x = my_tab, decreasing = TRUE)


```


You can quickly visualize the distribution of the `countries` vector: 

```{r, fig.align='center'}

sort_tab <- sort(x = my_tab, decreasing = TRUE)

barplot(sort_tab, ylab = "Counts", col = "steelblue")
```


# `search()` 


<center>

![from Unsplash by Ak Ka](https://images.unsplash.com/photo-1572965933956-65d960e6b11b?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=500&q=80)

</center>

`search()` is a nice little function that tells you which packages are attached to your current session. 

```{r}

search()

```


# `jitter()`

<center>

![from Unsplash by Dragisa Braunovic](https://images.unsplash.com/photo-1588416197225-c2554a33c536?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=750&q=80)

</center>


`jitter()` allows you to introduce some fluctuations to a vector of values


```{r}
# run mtcars$mpg to check the difference
jitter(mtcars$mpg)

```

# `comment()` 

<center>

![from Unsplash by Wolfgang Hasselmann](https://images.unsplash.com/photo-1545320059-b95959ddeb55?ixlib=rb-1.2.1&auto=format&fit=crop&w=752&q=80)

</center>

The `comment()` function is particularly useful when you want to bind some comments to a specific object. When the object is printed, the comments won't be displayed. 


```{r}

comment(mtcars) <- "This data frame has no NAs, go ahead !"

comment(mtcars)
```

The `attributes()` function will also retrieve the comments

```{r}

attributes(mtcars)


```

# `make.unique()` 

<center>

![from Unsplash by Wolfgang Hasselmann](https://images.unsplash.com/photo-1545481780-acff2f394ef6?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=752&q=80)

</center>



`make.unique()` is a pretty powerful function. It appends a sequence of numbers to duplicates in order to make vector's elements unique:


```{r}

countries <- c("Algeria", "Morocco", "Algeria", "Algeria", "Morocco", "Tunisia", "Morocco", "Tunisia")

make.unique(names = countries, sep = " -_- ")

```

# `startsWith()` and `endsWith()` 



<center>

![from Unsplash by Gilberto Olimpio](https://images.unsplash.com/photo-1580636938889-2df022d328d2?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=500&q=80)

</center>


`startsWith()`/`endsWith()` detect the elements of a vector (character) that start/end with a specific character(s):


```{r}

countries <- c("Armania", "Argentina", "Antalya", "Adelaide", "Abidjan")

startsWith(x = countries,
           prefix = "Ar")
```

```{r}

countries <- c("Armania", "Argentina", "Antalya", "Adelaide", "Abidjan")

endsWith(x = countries,
         suffix = "an")
```



# `quarters.Date()` 



<center>

![from Unsplash by Annie Spratt](https://images.unsplash.com/photo-1551163829-d8b73b2b0b94?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=667&q=80)

</center>


`quarters.Date()` converts a date to its corresponding quarter (Q1, Q2, Q3 or Q4): 



```{r}

my_dates <- c("2020-01-01", "2005-03-25", "2010-04-02", "2020-12-10", "2011-08-15")

quarters.Date(my_dates)

```



