---
title: Intro to the case_when function
author: 'Mohamed El Fodil Ihaddaden '
date: '2019-08-31'
tags:
  - tidyverse
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
lastmod: '2019-08-31T02:52:43+02:00'
slug: case-when-function
---

`case_when()` is a **dplyr** function that returns a value according to predifined conditions. It's a very powerful function though not very famous. In our example, we'll use the mtcars dataset (just as usual). 


```{r}
head(mtcars)
```

Suppose in the context of a study we want to categorize the cars into two factors : **Climate_Friendly** and **Climate_Unfriendly** according to the *horsepower (hp)* median. We can create a new column that incoporates this information using the `case_when()` and `mutate` functions:


```{r message=FALSE, warning=FALSE}
library(dplyr)
library(magrittr) # required for the %>% 

median <- median(mtcars$hp) # first, we calculate our median 

new_df <- mtcars %>% mutate(climat_categ = case_when(
    hp < median ~ "friendly",
    hp >= median ~ "unfriendly"
    ))


new_df$climat_categ <- as.factor(new_df$climat_categ)

new_df %>% select(hp, climat_categ)  # A quick check ! 
```


Let'us plot the count of the cars according to the 'climat_categ' variable. 

```{r}
library(ggplot2)

ggplot(new_df, aes(x =climat_categ))+ 
  geom_bar(stat = "count", fill = "darkolivegreen1", color = "blue") +
  theme_classic()+
  labs(title ="Count distribution of eco and non-eco friendly cars", x = "" )
  


```


