---
title: 'Importing and Visualizing Stock Market Data in R'
author: Mohamed El Fodil Ihaddaden
date: '2019-08-19'
slug: Importing-and-Visualizing-Stock-Market-Data-in-R
categories: []
tags:
  - Finance
  - Economics
subtitle: ''
summary: ''
authors: []
lastmod: '2019-08-19T16:36:05+02:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

```{r include=FALSE}
library(widgetframe)

```

In this tutorial, we'll analyze the French CAC 40 stocks. Listed CAC 40 firms can be found at the [Yahoo Finance website](https://finance.yahoo.com/quote/%5EFCHI/components/). 


## Getting the Symbols

First, we need to obtain the associated Yahoo symbols for each CAC 40 stock. Example: for Danone S.A, we have the correspnding symbol [BN.PA](https://finance.yahoo.com/quote/BN.PA?p=BN.PA). From Yahoo Finance we use the [DataPasta package](https://cran.r-project.org/web/packages/datapasta/vignettes/how-to-datapasta.html) to copy and paste the entire data table (We have modified the the column's name **% Change** to deal with R naming requirement). Then we select the **Symbol** column. 

```{r message=FALSE, warning=FALSE}

library(tidyverse)


base_data <- data.frame(stringsAsFactors=FALSE,
         Symbol = c("VIE.PA", "AI.PA", "BN.PA", "ACA.PA", "CA.PA", "ML.PA",
                    "BNP.PA", "RI.PA", "SU.PA", "VIV.PA", "ENGI.PA", "CAP.PA",
                    "LR.PA", "ORA.PA", "AC.PA", "HO.PA", "KER.PA", "GLE.PA",
                    "EN.PA", "SGO.PA", "SAN.PA", "DG.PA", "ATO.PA", "OR.PA", "SW.PA",
                    "UG.PA", "AIR.PA", "MC.PA", "FP.PA", "FTI.PA"),
   Company.Name = c("Veolia Environnement S.A.", "L'Air Liquide S.A.",
                    "Danone S.A.", "Crédit Agricole S.A.", "Carrefour SA",
                    "Compagnie Générale des Établissements Michelin",
                    "BNP Paribas SA", "Pernod Ricard SA", "Schneider Electric S.E.",
                    "Vivendi SA", "ENGIE SA", "Capgemini SE", "Legrand SA", "Orange S.A.",
                    "Accor SA", "Thales S.A.", "Kering SA",
                    "Société Générale Société anonyme", "Bouygues SA",
                    "Compagnie de Saint-Gobain S.A.", "Sanofi", "VINCI SA", "Atos SE", "L'Oréal S.A.",
                    "Sodexo S.A.", "Peugeot S.A.", "Airbus SE",
                    "LVMH Moët Hennessy - Louis Vuitton, Société Européenne", "TOTAL S.A.",
                    "TechnipFMC plc"),
     Last.Price = c(21.93, 123.35, 80.24, 10.26, 15.49, 93.32, 40.86, 158.05,
                    73.12, 25.64, 13.57, 106.35, 61.98, 13.6, 38.3, 102.9,
                    437.6, 22.35, 32.4, 32.33, 77.04, 96.82, 66.92, 236.7, 101.6,
                    19.34, 125.32, 355.2, 44.08, 21.73),
         Change = c(0, 0.15, 0.14, 0.03, 0.06, 0.48, 0.25, -1.05, 0.48, 0.17,
                    0.11, 1.05, 0.62, 0.14, 0.42, 1.2, 5.3, 0.28, 0.42, 0.43,
                    1.05, 1.34, 0.94, 3.4, 1.62, 0.34, 2.3, 6.75, 0.85, 0.75),
        Change_prc = c("0.00%", "+0.12%", "+0.17%", "+0.24%", "+0.39%", "+0.52%",
                    "+0.62%", "-0.66%", "+0.66%", "+0.67%", "+0.82%", "+1.00%",
                    "+1.01%", "+1.08%", "+1.11%", "+1.18%", "+1.23%", "+1.29%",
                    "+1.31%", "+1.35%", "+1.38%", "+1.40%", "+1.42%", "+1.46%",
                    "+1.62%", "+1.79%", "+1.87%", "+1.94%", "+1.97%", "+3.57%"),
         Volume = c(818027, 264260, 461968, 3363436, 772440, 275617, 3304556,
                    167013, 408078, 788907, 2077728, 221310, 258356, 3729428,
                    202482, 178755, 128660, 2969772, 264297, 640992, 1164010,
                    708090, 202814, 322746, 88158, 814983, 432166, 306987, 2264031,
                    777895)
)


Symbols <- base_data %>% select(Symbol)



```

# Loading Stock Prices Data

In order to import the stock prices, we use the **getSymbols()** function from the **quantmod** package. Let's say we are interested in the evolution of CAC 40 Adjusted Prices from the 15/07/2019 to 15/08/2019. The map(~Ad(get(.))) is used to extract the Adjusted Prices from the imported prices. 



```{r message=FALSE, warning=FALSE}

library(quantmod)

cac_prices <- getSymbols(Symbols$Symbol, 
                     src = "yahoo",
                     from = "2019-07-15",
                     to = "2019-08-15",
                     auto.assign = TRUE,
                     warnings = FALSE) %>% 
  map(~Ad(get(.))) %>% 
  reduce(merge) %>% 
  "colnames<-"(Symbols$Symbol)


head(cac_prices, n = 2) 

```


Using the getSymbols() function, we've obtained an **xts** object. xts or Extensible Time Series are dataframes indexed with dates. 



# Stock Prices Visualization with highcharter

**highcharter** offers highly reactive time series visualizations. Suppose we are interested in plotting French banks adjusted prices. 

```{r message=FALSE, warning=FALSE}
library(highcharter)


plot1 <- highchart(type = "stock") %>% 
  hc_title(text = "Stocks' Evolution of French CAC 40 listed banks") %>% 
  hc_add_series(cac_prices$ACA.PA, name="Crédit Agricole") %>% 
  hc_add_series(cac_prices$BNP.PA, name="BNP Paribas") %>% 
  hc_add_series(cac_prices$GLE.PA, name="Société Générale")
  

```

```{r echo=FALSE}
frameWidget(plot1)  
```



From the above chart, we observe a very small variation. This is not suprising as the the considered period is short (1 month). Let us consider a longer period (1 year). 
s

```{r}

Symbol_banks <- c("ACA.PA", "BNP.PA", "GLE.PA")

cac_banks <- getSymbols(Symbol_banks, 
                     src = "yahoo",
                     from = "2018-08-18",
                     to = "2019-08-18",
                     auto.assign = TRUE,
                     warnings = FALSE) %>% 
  map(~Ad(get(.))) %>% 
  reduce(merge) %>% 
  "colnames<-"(Symbol_banks)


plot2 <- highchart(type = "stock") %>% 
  hc_title(text = "Stocks' Evolution of French CAC 40 Listed Banks") %>% 
  hc_add_series(cac_banks$ACA.PA, name="Crédit Agricole") %>% 
  hc_add_series(cac_banks$BNP.PA, name="BNP Paribas") %>% 
  hc_add_series(cac_banks$GLE.PA, name="Société Générale")

```


```{r echo = FALSE}
frameWidget(plot2) 

```









