---
title: How to split a dataframe in R?
author: Mohamed El Fodil Ihaddaden
date: '2019-11-11'
slug: how-to-split-a-dataframe-in-R
categories: []
tags:
  - base R
  - R
subtitle: ''
summary: ''
authors: []
lastmod: '2019-11-11T22:18:06+01:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---


In this tutorial we'll work with the **ChickWeight** data. This database exhibits information on chickens' weights according to 4 different diets. 


```{r}

head(ChickWeight)

```

Suppose, we want to split our overal dataframe according to each type of diet. There is a simple function called `split()` that allows us to do that. **Note that the splitting argument must be a factor object.** Let's check this condition. 

```{r}

class(ChickWeight$Diet)

```


Indeed, the Diet variable is a factor. Now, we split our dataframe : 

```{r}

splitted_data <- split(ChickWeight, ChickWeight$Diet)

```

In this context, we can assign a name to each splitted data:

```{r}

Diet1 <- splitted_data$`1`  # The 1, 2, 3, 4 represent the diet factor levels

Diet2 <- splitted_data$`2`

Diet3 <- splitted_data$`3`

Diet4 <- splitted_data$`4`


```

Finally, let's check our result by printig the first values of Diet 3 and Diet 4: 

```{r}

head(Diet3)


```

```{r}

head(Diet4)


```


Great, it's perfectly working. 

