---
title: 'Three ways to get a summary of your data '
author: Mohamed El Fodil Ihaddaden
date: '2019-11-11'
slug: three-ways-to-get-a-summary-of-your-data
categories: []
tags:
  - R
subtitle: ''
summary: ''
authors: []
lastmod: '2019-11-11T19:02:15+01:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

## Summary 

To get a first idea about a certain dataframe's statistical metrics, there is of course the well known `summary()` function : 

```{r}

summary(mtcars)


```


However, there are other interesting functions that provide more structured and exhaustive information. 

## describe

The **psych** package has a function called `describe()` that provides several statistics that are not available within the `summary()` function. 


```{r}
library(psych)

describe(mtcars)

```


You can see that there is a star next to the *am* variable. This way, the `describe()` function is telling us that the *am* variable is a factor and it doesn't make sense to measure for example its mean or its standard deviation. 

In this context, there is an argument that allows us to omit non-numeric variables. 

```{r}

describe(mtcars, omit = TRUE)


```


## skim 


Finally, we present the `skim()` function from the **skimr** package which provides some statistics and a tiny nice histogram for the numerical variables (the histograms cannot be rendered in the website, but just try it out) !!! 

```{r}
library(skimr)

skim(mtcars)


```


I'm sure there may be many other super cool functions, please DM me at **@IhaddadenFodil** if I've missed something. 
